SIG Proceedings Paper in LaTeX Format

### 2019-09-27 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/conclusion.tex
changed back to paper without word freqs

This work provides an overview of the Skip Gram Model with negative Sampling (SGNS) and the numerous successful attempts of optimizing the throughput of the model. As this is the case, no effort went into optimizng the convergence time of the SGNS, therefore this work focused on this task. We decided to use advanced optimizers and input shuffling to achieve a better convergence time.This work proposes a slighlty altered version of the SGNS, where the idea is to compute the loss over the sum of a high number of training samples, i.e 2000,  instead of computing it for each individually.  We did this as it allowed us to compute more models and analyze the convergence time faster. We used the text8 dataset to train our model and  word similarity as a quality measure for the word embeddings (WE). We used the sate of the art implementation Gensim to compare our self. We did achieve a better convergence time than Gensim with Adam as an optimizer and the use of input shuffling. Gensim convereged in 4 epochs to a word similarity of 0.66 and our model only took 2 epochs to achieve the same quality. Those results still need to be confirmed with more datasets. Finally, if this work would be combined with an optimized throughput it  could improve the state of the art run time of the SGNS.

>>>>

This work provides an overview of the Skip Gram Model with negative Sampling (SGNS) and the numerous successful attempts of optimizing the throughput of the model. As this is the case, no effort went into optimizng the convergence time of the SGNS, therefore this work focused on this task. We decided to use advanced optimizers and input shuffling to achieve a better convergence time. This work proposes a slighlty altered version of the SGNS, where the idea is to compute the loss over the sum of a high number of training samples, i.e 2000,  instead of computing it for each individually. The batched version allowed us to compute more models and analyze the convergence time faster. We used the text8 dataset to train our model and  word similarity as a quality measure for the word embeddings (WE).
To compare our work, Gensim was used, which holds a state of the art implementation of the SGNS. We did achieve a better convergence time than Gensim with Adam as an optimizer and the use of input shuffling. Gensim convereged in 4 epochs to a word similarity of 0.66 and our model only took 2 epochs to achieve the same quality. Those results still need to be confirmed with more datasets. Finally, if this work would be combined with an optimized throughput it  could improve the state of the art run time of the SGNS.

### 2019-09-27 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/contribution.tex
changed back to paper without word freqs

\section{Batched Skip-Gram Model} \label{sec:contribution}
This section will give an overview of our implementation. The proposed implementation is a slighlty altered version of the origina Skip-Gram Model with negative sampling (SGNS). The idea behind this altering, is to compute the loss for multiple words and context pairs at the same time, this is inspired from mini-batch training, a model often used in other machine learning problem. Instead of computing the loss function of each pair, we computed the sum over a certain number of pairs (i.e 2000) as a loss. The exact process will be described in the following paragraph.
\subsection{Forwarding}
As the computation of the new loss function represents the challenging part of our implementation the forwarding method is explained step by step. Each time step is illustrated to make the explanation clearer.\\
Let $X = {(v_1,c_1),(v_2,c_2),(v_3,c_3)}$, be the training batch, where $(v_i,c_i)$ is a training sample constituted by a word and one of its context words. 
\subsubsection{Input}
The forwarding method will accept two vectors $v$ and $c$, and a matrix $A$ as an input. The first vector represents all the center words in a batch, the second one the context words. The Matrix represents the negative samples. The two vectors are of the same length, defined as $n$. The matrix must be of dimension $n \times k$, with $k$ being the number of negative samples per pair. This means the $i^{th}$ row will store the negative samples for the $i^{th}$ word context pair.\\
The input can be illustrated as follows: \\
$v = \begin{bmatrix}
v_1 & v_2 & v_3
\end{bmatrix}, c = \begin{bmatrix}
c1\\
c2\\
c3\end{bmatrix}$ and $A =
\begin{bmatrix}
k_{1,1} & k_{2,1} & k_{3,1}\\
k_{1,2} & k_{2,2} & k_{3,2}\\
k_{1,3} & k_{2,3} & k_{3,3}\\
\end{bmatrix}$\\

@@

This finishes the description of our batched approach.

\subsection{Distribution of word frequencies}
\subsubsection{Boxplot}
The quality of a batch is dependent on  the number of multiple appearances of words in the batch. The more frequently a word appears in a batch, the training quality of the word decreases. For this matter we chose to analyze the distribution of word frequencies. A box plot of the distribution, on the text8 dataset, can be found in Figure \ref{fig:boxplot_freq}. 
\begin{figure}[h]
\centering
\includegraphics[scale=0.3]{images/no_sampling_boxplot}
\caption{Distribution of word frequencies in the text8 datasedt}
\label{fig:boxplot_freq}
\end{figure} 
This shows that very few words appear very frequently, and are therefore responsible for the building of the most context pairs. For example the word \textit{the} creates 50\% of the pairs in the text8 dataset. In consequence if we create arbitrary batches, most words  will appear multiple times in a batch and the learning will be hindered. A solution would be to only allow words to appear once in a batch, in the next section we present such a solution with a variable batch size. \subsubsection{Batches}
To solve this problem we only chose to block multiple appearance in a batch on the center words. The context words can be the same for two different center words.  Each batch will contain the highest possible number of pairs where each center word does not appear twice. The problem with this approach is that the batch size decreases very rapidly. This is due to the distribution of words. In consequence the average batch size is small and training time is increased, to remedy this problem we tried to reduce the number of very small batches, while maintaining the quality of learning. 
\subsubsection{Deletion of outliers}
To solve this matter, we looked again at the distribution of word frequencies (Figure \ref{fig:boxplot_freq}). We experimented with different settings (mean, median, average, outlier) and found that deleting all words that had a frequency greater than the 99\% was best. 
This technique drastically reduces the dataset size.

The next section will describe our experimental setting and the possibilities that exists to evaluate the quality of word embeddings.

>>>>

\section{Batched SkipGramModel} \label{sec:contribution}
This section will give an overview of our implementation. The proposed implementation is a slighlty altered version of the origina Skip-Gram Model with negative sampling (SGNS). The idea behind this altering, is to compute the loss for multiple words and context pairs at the same time, this is inspired from mini-batch training, a model often used in other machine learning problem. Instead of computing the loss function of each pair, we computed the sum over a certain number of pairs (i.e 2000) as a loss. The exact process will be described in the following paragraph.
\subsection{Forwarding}
As the computation of the new loss function represents the challenging part of our implementation the forwarding method is explained step by step. Each time step is illustrated to make the explanation clearer.\\
Let $X = {(v_1,c_1),(v_2,c_2),(v_3,c_3)}$, be the training batch, where $(v_i,c_i)$ is a training sample constituted by a word and one of its context words. 
\subsubsection{Input}
The forwarding method will accept two vectors $v$ and $c$, and a matrix $A$ as an input. The first vector represents all the center words in a batch, the second one the context words. The Matrix represents the negative samples. The two vectors are of the same length, defined as $n$. The matrix must be of dimension $n \times k$, with $k$ being the number of negative samples per pair. This means the $i^{th}$ row will store the negative samples for the $i^{th}$ word context pair.\\
The input can be illustrated as follows: \\
$v = \begin{bmatrix}
v_1 & v_2 & v_3
\end{bmatrix}, c = \begin{bmatrix}
c1\\
c2\\
c3\end{bmatrix}$ and $A =
\begin{bmatrix}
k_{1,1} & k_{2,1} & k_{3,1}\\
k_{1,2} & k_{2,2} & k_{3,2}\\
k_{1,3} & k_{2,3} & k_{3,3}\\
\end{bmatrix}$\\

@@

This finishes the description of our batched approach. The next section will describe our experimental setting and the possibilities that exists to evaluate the quality of word embeddings.
